A continuous urbanization index to assess phylogenetic adaptations in urban bird living

### 2018-05-15 ms.Rmd
combining the 'phylo' para and the 'brain size' para into one

# ABSTRACT
Understanding species' responses to urbanization is critical to the conservation of urban biodiversity. Previous work assessing ecological and life history traits that are associated with urban birds has elucidated differing and contrasting patterns for a wide variety of traits. Much of this work has relied on a priori grouping of species into discrete groups (e.g., urban avoiders, exploiters, and adapters or even binary presence/absence in urban areas), and has been spatially and/or temporally limited. Here, we provide a novel and easily adapted methodology which uses a continuous urbanization index to assess which ecological and life history traits most influence a species' ability to be found in urban environments. A continuous measure of urbanization is critical as it fully accounts for the inherent differences in species' responses to urban environments. We further expand on previous studies by incorporating a total of 477 species in our analyses, a number far greater than similar studies, to the best of our knowledge. We find a significant phylogenetic effect of our urbanization index and thus account for the non-linear relationships among species in our modelling. Niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors of urbanization. We also find that residual brain mass is a critical predictor within our phylogenetic models, suggesting that those species within a clade which have the largest relative brain size are most readily adapting to urban development. Urbanization is selecting against specialists, granivores, and insectivores. These results have important implications for aiding the preservation and persistence of urban avian biodiversity.

@@

Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index.  This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertaintly. Conversely, specilized feeding strategies were negatively associated with the urbanization index: granivore and insectivore specialists were much less likely to be found in urban environments. These results suggest that urbanization is, in general, excluding most specialist species.  SOMETHING ABOUT IMPLICATIONS FOR CONSERVATION / FRAGMENTATION.

The non-random distribution of the urbanization index across the phylogeny (Table 1) is interesting.  This result contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes].  Interestingly many of the predictor traits---generalism, clutch size---have even stronger phylogenetic signal, so it's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny.  It is worth noting that this results is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @major2010museum].

The effect of treating species as non-independent by considering phylogeny in the context of the linear model was apparent. For example, there was an increase in parameter estimates of the most important parameters (e.g., clutch size and feeding habitat generalism; Figures 4 and 5), and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig. S2). Further, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass and largest body size are most suited to surviving in urban areas [@major2010museum]. Conversely, other predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural). For the variables which became less important, it is likely that there may be particular clades that include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence].

Brain size at a given body size is an especially interesting trait--it's becomes more important in the phylogenetic model compared to the non-phylogenetic model.  While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], we confirm those which do support the hypothesis that brain size is heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain]. Because of the phylogenetic signal in both brain size and urbanization, it is crutial to investigate this relationship in a phylogenetically informed manner.

Certian aspects of the trait-urbanization relationship appear to consistent across the globe. CHANGE PHRASING BELOW HERE: The relationship with annual fecundity and urbanization in birds is consistent with work in Europe  [@croci2008does; @moller2009successful]. The that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes]. While others have found that urbanization supports granivorous species [@chace2006urban; @kark2007living; @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden]. We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].

>>>>

# ABSTRACT
Understanding the suite of ecological and life history traits which are most associated with urbanization enables a predictive ability to understand the species which are readily suited to adapt with our increasingly urbanizing world. Previous work investigating this relationship has found contrasting patterns for a specific trait's association with urbanization. Here, we provide a novel and scalable methodology relying on a continuous urbanization index to assess ecological and life history traits which most influence a species' ability to be found in urban environments. In a continental-scale analysis, incorporating 477 species, we find that niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors of urbanization. Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urban environments. Our results suggest that the enhancement and persistence of avian biodiversity in urban areas relies on developing suitable habitat to serve a wide range of life history strategies.

@@

Niche breadth, as measured by different types of generalism, was significantly associated with our urbanization index. This result was robust across non-phylogenetic and phylogenetic models, model averaging approaches, and across phylogenetic uncertainty. Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments. These results suggest that urbanization is, in general, excluding most specialist species.  SOMETHING ABOUT IMPLICATIONS FOR CONSERVATION / FRAGMENTATION.

The non-random distribution of the urbanization index across the phylogeny (Table 1) contrasts with previous studies [@evans2011makes], but this contrasting finding is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes]. It's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny and we demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear model. For example, some predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural). For these variables, it is likely that there may be particular clades which include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence]. Conversely, some of the most important predictor variables (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization when phylogeny was accounted for, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig. S2). Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas. While some studies have found a lack of evidence that relative brain size is critical in urban birds [@kark2007living; @evans2011makes], our results suggest that brain size is indeed heavily correlated with urban birds [@moller2009successful; @maklakov2011brains; @moller2015brain]. This finding highlights the necessity to investigate these relationships in a phylogenetically informed manner. Lastly, it is worth noting that these results are consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation; @major2010museum].

Certian aspects of the trait-urbanization relationship appear to be consistent across the globe. CHANGE PHRASING BELOW HERE: The relationship with annual fecundity and urbanization in birds is consistent with work in Europe  [@croci2008does; @moller2009successful]. The that urbanization appears to select against insectivorous species [@chace2006urban; @evans2011makes]. While others have found that urbanization supports granivorous species [@chace2006urban; @kark2007living; @evans2011makes], we found the opposite trend, but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden]. We did not find any strong support for our hypotheses on different nesting types, but we did find that in general, species which tended to be more social were at an advantage to city living, confirming previous studies [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living].
